### 2015-10-05 response.tex
Answered first reviewer

\documentclass[letterpaper]{article}
\usepackage{url}
\usepackage[margin=1in]{geometry}
\newcommand{\breview}{\begin{quotation}\begin{bf}\noindent}
\newcommand{\ereview}{\end{bf}\end{quotation}}

\begin{document}

\section*{Reviewer \#1}

\breview

The authors give a good overview of the existing methods to predict the retention time.  However, the description of the methods is sometimes too brief, and slightly biased towards their own developed methods. Extending the descriptions of the methods that are not developed by the authors should be considered. The applications and discussion section are an added value to the paper.

\ereview

We would like to thank the reviewer for the positive feedback. We have revised the sections were the other methods contribution is discussed

\breview
Additional comments:
\begin{itemize}
\item Legend of Figure 2 overlaps with text on page 3
\item Table 1 is not fully visible; additionally in the table “Elude” is written, while in the text “ELUDE” is used
\end{itemize}
\ereview
We have corrected this.
\breview
\begin{itemize}

\item The authors write on page 5 and 6: ``The algorithm reported good
accuracy, especially taking into account that it only used information
about the amino acid composition of the peptides.'' This statement
seems to be contrasting with the following comment found on page 3:
``As a consequence, given a chromatography setup, the sequence of a
peptide should provide virtually all the information needed to predict
its retention time''. To the latter statement several other
contrasting statements can be found on the same page.
\end{itemize}
\ereview
We added a qualifying paragraph to the later statement.
``As a consequence, given a chromatography setup, the sequence of a
peptide should provide virtually all the information needed to predict
its retention time. However, in practice this is not as straight
forward as it might seem, as it for instance is difficult to predict a
peptide's three dimensional structure from its amino acid sequence.''
\breview
\begin{itemize}
\item The paragraph in Section 4 on the modified ELUDE is not completely clear. An extended discussion would be useful
\end{itemize}
\ereview

We revised this section to better explain how we incorporated
retention time predictions of modified peptides.

\breview
Typos:
\begin{itemize}
\item Page 6: Klammer et al. $\to$ Klammer {\em et al.}
\item Page 6: Moruz et al.[22] $\to$ Moruz {\em et al.} [22]
\item Page 7: Strittmatter et al. $\to$ Strittmatter {\em et al.}
\item Page 7: Klammer et al. $\to$ Klammer {\em et al.}
\item Page 8: Krokhin et al. $\to$ Krokhin {\em et al.}
\item Page 8: bigdata $\to$ big data
\item Page 9: retention time[57] $\to$ retention time [57].
\item Page 9: DIA $\to$ data independent acquisition
\end{itemize}

\ereview

Thank you for highlighting these typos. We have now corrected them.

\section*{Reviewer \#2}

\breview
\ereview

\end{document}